---
title: "InterimAnalysis"
author: "Zoe Woodhead"
date: "21/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(yarrr)
require(kableExtra)
```

## Demographics

This is an interim analysis of fTCD data for the COLA registered report project.

```{r readdata, echo=FALSE, warning=FALSE}
alldata <- read.csv('fTCD_data.csv')

tasks <- c('WG','SG','PD','WC','SC','SD')
sites <- c('Bangor','Lancaster','Lincoln','Oxford','UCL','UWA')
alldata$Site <- as.factor(alldata$Site)
alldata$Sex_F <- as.factor(alldata$Sex_F)
levels(alldata$Sex_F) <- c('Male', 'Female')
alldata$Hand_R <- as.factor(alldata$Hand_R)
levels(alldata$Hand_R) <- c('Left', 'Right')
alldata$Gorilla_ID <- as.factor(alldata$Gorilla_ID)

# How many participants were tested?
n_attended <- length(alldata$Gorilla_ID)

# How many datasets were acquired?
n_acquired <- length(na.omit(alldata$A_nMark))

# Participant Counts
ParticipantCounts <- matrix(nrow=6, ncol=4, data=NA)
colnames(ParticipantCounts) <- c('M_LH', 'M_RH','F_LH', 'F_RH')
rownames(ParticipantCounts) <- sites

for (s in 1:6){
  site_data <- alldata %>% filter(Site == sites[s])
  mytable <- table(site_data$Sex_F, site_data$Hand_R)
  ParticipantCounts[s, ] <- c(mytable[1,], mytable[2,])
} 
kable(ParticipantCounts)

# Omit rows with no fTCD data
mydata <- alldata %>% filter(!is.na(A_nMark))

# Sex and handedness summary stats
n_female_L <- length(which(mydata$Sex_F=='Female' & mydata$Hand_R=='Left'))
n_female_R <- length(which(mydata$Sex_F=='Female' & mydata$Hand_R=='Right'))
n_male_L <- length(which(mydata$Sex_F=='Male' & mydata$Hand_R=='Left'))
n_male_R <- length(which(mydata$Sex_F=='Male' & mydata$Hand_R=='Right'))


```

In total, `r n_attended` participants attended an fTCD testing session, and data was successfully acquired from `r n_acquired` of those participants; the failure rate due to lack of a clear Doppler signal was `r round((n_attended - n_acquired)/ n_attended *100, 2)`%.  

Of the participants from whom data was successfully acquired, `r round((n_female_L + n_female_R) / n_acquired * 100, 2)`% were female and `r round((n_female_L + n_male_L) / n_acquired * 100, 2)`% were left handed.


## Data Quality and Outliers

``` {r dataqual, echo=FALSE, warning=FALSE}
# Identify outliers
for (t in 1:6){
  SEcol <- which(colnames(mydata) == paste0(LETTERS[t], '_mean_se'))
  Q3<-quantile(mydata[ , SEcol],.75,na.rm=TRUE)
  Q1<-quantile(mydata[ , SEcol],.25,na.rm=TRUE)
  Qlimit<-Q3+2.2*(Q3-Q1)
  
  # If there are at least 10 trials, include the data
  excludecol = which(colnames(mydata) == paste0(LETTERS[t], '_exclude'))
  trialscol = which(colnames(mydata) == paste0(LETTERS[t], '_N'))
  mydata[which(mydata[ , trialscol] > 9), excludecol] <- 0
  
  # If the SE is too high, exclude the data
  mydata[which(mydata[ , SEcol] > Qlimit) , excludecol] <- 1
}

# Count number of missing or excluded datapoints per task
n_excludeLI = matrix(data=NA, nrow=1, ncol=6)
for (t in 1:6){
  excludecol = which(colnames(mydata) == paste0(LETTERS[t], '_exclude'))
  n_excludeLI[t] = length(which(mydata[ , excludecol] > 0)) + length(which(is.na(mydata[ , excludecol])))
}

# Should any participants be excluded?
mydata$Exclude <- 0
tmp <- mydata$A_exclude + mydata$B_exclude + mydata$C_exclude + mydata$D_exclude + mydata$E_exclude + mydata$F_exclude
mydata$Exclude[which(tmp > 1)] = 1
n_excluded = length(which(mydata$Exclude == 1))

# Drop these participants from subsequent analysis
mydata <- mydata %>% filter(Exclude == 0)

```

Laterality Indices (LIs) were calculated for all six tasks for all participants. LIs were omitted if the data quality was poor, e.g. if there were insufficient usable trials, or if the standard error across the trial-by-trial LI values was too high (using the Hoaglin Iglewicz criterion). In total, `r sum(n_excludeLI)` LI values (`r round(sum(n_excludeLI) / (6*n_acquired) *100, 2)`%) were excluded from the analysis. If an individual participant had more than one LI value excluded, all of their data was excluded from futher analysis; this was the case for `r n_excluded` partipants, leaving `r n_acquired - n_excluded` remaining. 

Split half reliability for each task was assessed by calculating laterality indices using odd or even trials only, and taking the correlation between the two. The table below shows these split half correlations for each task.

```{r reliability, echo=FALSE, warning=FALSE}
# Split-Half Reliability

SHdata <- mydata %>% 
  select(Gorilla_ID, Hand_R, A_mean_odd, A_mean_even, B_mean_odd, B_mean_even, 
         C_mean_odd, C_mean_even, D_mean_odd, D_mean_even, 
         E_mean_odd, E_mean_even, F_mean_odd, F_mean_even)

# Test-retest comparisons of LI values
LI_splithalf <- data.frame(matrix(NA, nrow=6, ncol=3))
colnames(LI_splithalf)<-c("Task", "r", "p")

LI_splithalf$Task <- LETTERS[1:6]

# Loop through tasks
for (t in 1:6){
  # Calculate correlation
  my_test <- cor.test(SHdata[ , (2*t+1)],    # Odd trials
                      SHdata[ , (2*t+2)])   # Even trials
  # Organise output
  LI_splithalf$r[t] <- round(my_test$estimate, 2)
  LI_splithalf$p[t] <- round(my_test$p.value, 8)}
kable(LI_splithalf)

```



## LI Summary Statistics

The pirate plot shows LI values for the six tasks (A = Word Generation, B = Sentence Generation, C = Phonological Decision, D = Word Comprehension, E = Sentence Comprehension and F = Syntactic Decision) for left and right handed participants. One sample t-tests were computed to test whether the group LI values differed significantly from zero, i.e. showed significant lateralisation. In left handers, tasks A, B and C were left lateralised; task D was right lateralised; and tasks E and F were not significantly lateralised. In right handers, all six tasks differed significantly for zero. (NB - no correction for multiple comparisons). Between-group t-tests were also computed to test whether lateralisation differed between left and right handers. In all cases, LI values were significantly stronger in the right handers than the left handers.

```{r LIstats, echo=FALSE, warning=FALSE}

LIdata <- mydata %>% 
  select(Gorilla_ID, Hand_R, A_mean_LI, B_mean_LI, C_mean_LI, D_mean_LI, E_mean_LI, F_mean_LI)
colnames(LIdata) <- c('ID','Hand','A','B','C','D','E','F')
longdata <- pivot_longer(data = LIdata, cols = c(3:8), names_to = 'Task', values_to = 'LI')
pirateplot(data = longdata, LI ~ Task * Hand)
abline(h=0)
title(main=paste0('Pirate Plot of LI Data, n=',length(LIdata$ID)))

# One sample t-tests: Is LI different to zero?
t1_results = matrix(data=NA, nrow= 12, ncol = 4)
t1_results = as.data.frame(t1_results)
colnames(t1_results) = c('Task','Hand','t','p')
hand = c('Left','Right')
for (h in 1:2){
  for (t in 1:6){
    myt = t.test(LIdata[which(LIdata$Hand == hand[h]), t+2])
    t1_results$Task[6*(h-1)+t] = LETTERS[t]
    t1_results$Hand[6*(h-1)+t] = hand[h]
    t1_results$t[6*(h-1)+t] = round(myt$statistic, 3)
    t1_results$p[6*(h-1)+t] = round(myt$p.value, 3)
  }
}
kable(t1_results)

# Two sample t-tests: Do left and right handers differ?
t2_results = matrix(data=NA, nrow= 6, ncol = 3)
t2_results = as.data.frame(t2_results)
colnames(t2_results) = c('Task','t','p')
for (t in 1:6){
    myt = t.test(LIdata[which(LIdata$Hand == 'Left'), t+2], 
                 LIdata[which(LIdata$Hand == 'Right'), t+2], paired = FALSE, )
    t2_results$Task[t] = LETTERS[t]
    t2_results$t[t] = round(myt$statistic, 3)
    t2_results$p[t] = round(myt$p.value, 3)
}
kable(t2_results)

```



